18.821 Project 2

### 2014-10-27 paper/paper.tex
added introduction

\section{Terms in the sequence $s_n(312)$}

\begin{figure}
\begin{tikzpicture}[level distance=0.5cm, growth parent anchor={south}, nodes={anchor=north}]
\tikzstyle{hollow node}=[circle,draw,inner sep=0.1mm, text width=1.36cm,align=center]
\tikzstyle{level 1}=[sibling distance=6cm] 
\tikzstyle{level 2}=[sibling distance=3cm] 
\tikzstyle{level 3}=[sibling distance=1.8cm]
\tikzstyle{level 4}=[sibling distance=0.9cm]
\node [hollow node]{(1)}
	child {
		node[hollow node] {(1,2)}
		child{
			node[hollow node]{(1,2,3)}
			child{node[hollow node]{(1,2,3,4)}}
			child{node[hollow node]{(1,2,4,3)}}
			child{node[hollow node]{(1,4,2,3)}}
			child{node[hollow node]{(4,1,2,3)}}
		}
		child{
			node[hollow node]{(1,3,2)}
		}
	}
    	child {
		node[hollow node] {(2,1)}
    		child {
			node[hollow node] {(2,1,3)}
			child{node{\vdots}}
		}
      		child {
			node[hollow node] {(2,3,1)}
			child{node{\vdots}}
		}
		child {
			node[hollow node] {(3,2,1)}
			child{node{\vdots}}
		}
    	};

@@

@@

\section{Avoidance of permutations of $T_n$}
We now impose further restrictions on the set of permutations $S_n$. This new set of permutations, $T_n$ is defined follows.

>>>>

\section{Introduction}

In this paper we will consider investigate pattern avoidance in permutations of finite sets of positive integers.

\begin{definition}
Two finite sequences $a_1,\cdots a_k$ and $b_1,\cdots b_k$ have the same relative order if the $i^\text{th}$ largest entry of each unordered set $\{a_1,\cdots,a_n\}$ and $\{b_1,\cdots, b_n\}$ appears in the same position in $a_1\cdots a_k$ and $b_1\cdots b_k$ for each $1\le i\le k$. 
\end{definition}

\begin{definition}
A finite sequence of numbers $a_1a_2\cdots a_n$ avoids a sequence $b_1\cdots b_k$ with $n\ge k$ if no subsequence $a_{i_1}a_{i_2}\cdots a_{i_k}$ of $a_1a_2\cdots a_n$ has its terms in the same relative order as $b_1\cdots b_k$. 
\end{definition}

\begin{example}
The sequence 15432 avoids 123 while the sequence 13425 does not. The latter example fails because 134 appears in 15342 with the same relative order as 123. 
\end{example}

\begin{remark}
Let $S_n$ be the set of permutations of $\{1,\cdots, n\}$. There are two conventions commonly used to write elements of $S_n$. The first is $\emph{one-line}$ notation where permutations in $S_n$ are written explicitly, e.g. 43215 is a permutation of $\{1,2,3,4,5\}$ in $S_5$.

The second is \emph{cycle notation}, where a cycle $(a_1\cdots a_n)\in S_n$ with $1\le a_i\le n$ acts on a sequence by sending $a_1\mapsto a_2$, $a_2\mapsto a_3$, $\cdots$, $a_n\mapsto a_1$. For example, the cycle $(14)(23)$ sends $12345$ to $43215$. Unless otherwise stated, we will refer to permutations in one-line notation. 
\end{remark}

\begin{definition} 
When $n\ge k$, a permutation $\sigma\in S_n$ avoids a permutation $\pi\in S_k$ if $pi$ and $\sigma$ avoid each other when written as sequences in one-line notation. 
\end{definition}

\begin{definition}
For a permutation $\pi$, we write $s_n(\pi$ for the number of permutations in $S_n$ that avoid $\pi$. 
\end{definition}

In this paper we concern ourselves with the sequences $s_n(\pi)$ for various $\pi$. In section \ref{S3} we will consider sequence $s_n(\pi)$ for $\pi\in S_3$, and in section \ref{S4} we will consider sequence $s_n(\pi)$ for $\pi\in S_4$. In section \ref{Tn} we consider a variation on the avoidance problem that considers the problem of avoiding sequences in a subset $T_n$ of $S_n$.

\subsection{Generating Trees}

Before continuing we introduce generating trees, a tool we can use to study pattern avoidance. The generating tree for the sequence $s_n(\pi)$ is the infinite tree whose vertices are elements of $S_n$ that avoid $\pi$. To construct the tree, we define the children of an arbitrary node. Let $\omega\in S_k$ be a node of the generating tree of $s_n(\pi)$. Then the children of $\omega$ are elements of $S_{k+1}$ that avoid $\pi$ and that can be obtained by inserting the integer $k+1$ between integers in $\omega$. Figure \ref{fig:M1} is an example of the generating tree of $s_n(312)$.

\begin{figure}[h!]
\begin{tikzpicture}[level distance=0.5cm, growth parent anchor={south}, nodes={anchor=north}]
\tikzstyle{hollow node}=[circle,draw,inner sep=0.1mm, text width=1.36cm,align=center]
\tikzstyle{level 1}=[sibling distance=6cm] 
\tikzstyle{level 2}=[sibling distance=3cm] 
\tikzstyle{level 3}=[sibling distance=1.8cm]
\tikzstyle{level 4}=[sibling distance=0.9cm]
\node [hollow node]{(1)}
	child {
		node[hollow node] {(1,2)}
		child{
			node[hollow node]{(1,2,3)}
			child{node[hollow node]{(1,2,3,4)}}
			child{node[hollow node]{(1,2,4,3)}}
			child{node[hollow node]{(1,4,2,3)}}
			child{node[hollow node]{(4,1,2,3)}}
		}
		child{
			node[hollow node]{(1,3,2)}
		}
	}
    	child {
		node[hollow node] {(2,1)}
    		child {
			node[hollow node] {(2,1,3)}
			child{node{\vdots}}
		}
      		child {
			node[hollow node] {(2,3,1)}
			child{node{\vdots}}
		}
		child {
			node[hollow node] {(3,2,1)}
			child{node{\vdots}}
		}
    	};

@@

\section{Terms in the sequence $s_n(312)$}
\label{S3}

@@

\section{Conjectures on $s_n(\pi)$ for $\pi\in S_4$}
\label{S4}

\section{Avoidance of permutations of $T_n$}
\label{Tn}
We now impose further restrictions on the set of permutations $S_n$. This new set of permutations, $T_n$ is defined follows.
